### 2025-11-04 Paper/paperdraft_fall2025.tex
Updates from Overleaf

    In this paper, I study the effect of specialized executive training on firm behavior when the firm experiences an external shock to their objective function. I do this specifically in the context of US hospitals for two reasons. First, in the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs, as the US preforms poorly in these metrics compared to other similar countries. Yet, little is known about how leaders affect health care value in the hospitals they run. Second, nonprofit hospitals are particularly important in this context, and it is unclear whether existing findings on executive characteristics can be extrapolated to firms with presumably different incentives and objectives than those which have been previously studied.

@@

    The first prediction of the model is that, all else held equal, hospitals that value profit will have a lower quality than those who value patient welfare in a fee-for-service setting. Second, a hospital's change in quality once pay-for-performance is enacted depends on how much weight the hospital places on profit vs. patient welfare. Particularly, hospitals with higher weight on profit respond more than hospitals with higher weight on patient welfare. Therefore, factors that affect weight $\alpha$ will be correlated with choice of quality, and with how quality changes under new pay-for-performance incentives. In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$.

@@

    I use several data sets to analyze the relationship between executives with clinical training and hospital response to financial incentives on quality. I construct novel information on individual hospital executives by scraping publicly available tax forms, and merge this data to publicly available data sets containing other hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data. Ultimately, I employ this hospital-level data spanning from 2010-2014, which contains executive team characteristics, mortality and readmission rates as measures of quality, and other characteristics. In this section, I describe the institutional setting and data collection process. A more detailed description of the construction of this data can be found in Appendix \ref{appendixdata}.

@@

    Nonprofit hospitals, as with most nonprofits, are typically governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams. Some hospital executives specialize in health care settings by earning a degree in health care management, or an MBA specific to health care. There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can also fill other executive positions. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

@@

    For each hospital in the tax form data, I merge to hospital matches in the AHA data based on name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optical character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990. While board members are an important aspect of hospital organization, I do not include them in this analysis as they are beyond the scope of this paper. I extract text spanning 2009-2015, but I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014. This yields 852 nonprofit hospitals that are matched to AHA data, and with complete text information in each year.

    Finally, I verify clinical training of those who are identified as doctors, and obtain additional information about clinically trained executives by matching names to the National Plan and Provider Enumeration System (NPPES) database of all registered physicians. I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}. In total, there are 813 clinically trained executives. On average, they are 52 years old, and 11\% of clinically trained executives are female. While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs. The most common specialty is internal medicine.

@@

    I focus on executive teams as a whole instead of specific positions because the organization of teams varies so drastically across hospitals. It is not uncommon for two hospitals to have the same position but the tasks performed are drastically different, or to have two distinct positions that essentially perform the same tasks. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.

    \subsection{Pay-for-Performance Policies Targeting Readmission and Mortality Rates}\label{sec:hrrp}

    The purpose of this paper is to investigate hospital behavior in the midst of financial incentives that change how hospitals receive payment for quality of care. Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP). HRRP focused on penalizing hospitals with poor quality, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.

@@

    In Table \ref{tab:sumstats_samples_stable}, I present means of relevant variables for several hospital samples used throughout the paper. First, in column (1), are means for hospitals who ever have a clinical executive at any point in the sample. In column (2) are means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period. The main analysis compares hospitals in columns (2) and (3), and supplemental analyses leverage hospitals in column (1).

@@

    Next, I plot weighted average readmission rates, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}. The effective start year of pay-for-performance incentives is shown as a dotted line in 2012. Average readmission rates trend similarly prior to 2012, and all hospital types exhibit a decrease in raw readmission rates after 2012. However, there is a faster rate of decrease for hospitals who have a clinically trained executive at some point, or who never have a clinically trained executive. When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat. Finally, hospitals that never have clinical executives consistently have less complex patients, but there is no observable change in raw case mix index over time.

    \begin{figure}[ht!]
    \centering
        \caption{Outcomes Over Time}
        \includegraphics[width=.8\textwidth]{Objects/outcomes_graph.pdf}
        \label{fig:outcomes_graph}
    \end{figure}

@@

    I test empirically whether having clinically trained executives matters in how hospitals respond to the pay-for-performance policies effective in 2012. The quality metrics targeted by the programs are readmission and mortality rates, so I use these as outcomes. I limit the sample to hospitals who always have clinical executives or never have clinical executives. That is, the hospitals in this analysis do not change their propensity to hire a clinical executive during the sample period.\footnote{An important assumption for this analysis is that this sample selection is not correlated with the program enactment. I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs. }

    I ultimately employ a synthetic difference-in-differences estimation strategy that mirrors two-way-fixed effects, where the key independent variable is the interaction between having clinically trained executives and being in years after pay-for-performance took effect. All hospitals had incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals. Thus, I leverage the timing of the policy change, rather than the policies themselves.

    One may be concerned about potential differences in quality trends between clinical and non-clinical executive team hospitals leading to a violation in the parallel trends assumption. This is why I implement a synthetic difference-in-differences estimation strategy (\cite{arkhangelsky2021synthetic}). In particular, I weight more heavily any non-clinical team hospitals with pre-trends similar to clinical team hospitals in readmission, mortality, and baseline case mix. Further, I more heavily weight time periods that balance pre- and post-program outcomes for non-clinical teams. The control group consists of approximately 240 hospitals, and I show the fairly uniform distribution of their weights in Appendix \ref{app:sumstats}.

    I present the unweighted two-way-fixed-effects specification equation for clarity, and the identification assumptions are conceptually similar for synthetic difference-in-differences. In Appendix \ref{app:otherestimations}, I show that the findings are robust to using two way fixed effects, matching, or synthetic control.

    \begin{equation}
    \label{eq:clinical}
    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

    \noindent The variable $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively.

    Once incorporating the weights used in the synthetic differences-in-differences method, we ensure parallel pre-trends in the outcome variables. Thus, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012. This includes an implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team. Finally, I assume that there is no anticipation of the policy change before 2012. The rules of the policies were announced in October of 2011, so 2012 was the first full year that hospitals could meaningfully respond to the policies.

    The estimated difference in readmission and mortality rates between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}. Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began. Panel \ref{fig:read_synth_clinical} shows the estimated difference in readmission rates, and reveals that non-clinical teams decrease readmissions by .3 (standard error .13) more than clinical teams. This is a relatively small magnitude given an average readmission rate of 21, but reveals a meaningful difference in behavior given that both hospital types significantly decrease readmissions. In Appendix \ref{app:condition}, I show that these results are primarily driven by pneumonia patients.

    In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality. While the magnitude of the difference is similar to that for readmission rates (relative to the mean), mortality is a noisy measure, and I cannot rule out null effects. I investigate whether the estimated difference in readmission rate response could be due to differential participation in selective patient practices in Appendix \ref{app:casemix}, and find that this is not the case.

@@

    I present the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}. In panel \ref{fig:belowmed_read_synth_clinical}, I compare executive teams with no clinical training to those with a fraction of clinically trained executives up to the median value, 20\%. With an average treatment effect of .21 (standard error .31), hospitals having this range of clinically trained executives are not driving the main finding. However, in panel \ref{fig:abovemed_read_synth_clinical}, I compare hospitals with no clinically trained executives to hospitals with a fraction of clinically trained executives over the median of 20\%. I find that hospitals without clinically trained executives decrease readmissions by .6 (standard error .14) more than hospitals with clinically trained executives. Thus, the readmission rate response is driven by having a higher proportion of clinical executives relative to no clinical executives.\footnote{This result holds even when limiting the sample to hospitals with a similar total executive team sizes (Appendix \ref{app:binned})}

@@

    \section{Are Clinical Executives Less Profit Driven?}\label{sec:forprofit}

    The hypothesis driven from the theory described in Section \ref{sec:theory} is that clinically trained executives put more weight on quality of care before the financial incentives on quality began. To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals. Then, I compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change. While I do not have information on leadership teams of for-profit hospitals, on average, they have revealed through their ownership status that they more heavily weigh profit. Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are. Thus, I estimate the weighted synthetic difference-in-differences version of

@@

\section{Signaling vs. Managing Decomposition}\label{sec:sig_man}

@@

    \subsection{Analysis of Executive Team Changes} \label{sec:changes}

@@

     \subsection{Estimation}\label{sec:decomp}

@@

    \section{Heterogeneity by Physician Specialty}

@@

@@

        \begin{figure}[ht!]
     \caption{Effect of Clinical Training on Readmission Rates, Binned Treatment}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho \leq \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_belowmedread_md_nomd_size_synth_graph.pdf}
         \label{fig:belowmed_read_synth_clinical}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho > \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_abovemedread_md_nomd_size_synth_graph.pdf}
         \label{fig:abovemed_read_synth_clinical}
     \end{subfigure}
        \label{fig:cont_clinicalsynthdid_read_size}
    \end{figure}

@@

            \begin{figure}[ht!]
     \caption{Effect of Clinical Training on Mortality Rates, Binned Treatment}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho \leq \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_belowmedmort_md_nomd_size_synth_graph.pdf}
         \label{fig:belowmed_read_synth_clinical}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho > \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_abovemedmort_md_nomd_size_synth_graph.pdf}
         \label{fig:abovemed_read_synth_clinical}
     \end{subfigure}
        \label{fig:cont_clinicalsynthdid_size}
    \end{figure}

>>>>

    In this paper, I study the effect of specialized executive training on firm behavior when the firm experiences an external shock to their objective function, specifically in the context of US hospitals. I leverage this setting for two reasons. First, in the US, health care makes up almost 20\% of spending (\cite{CMSNationalHealthExpenditure}), and a major goal of policymakers is to improve the quality of health care without increasing costs, as the US preforms poorly in these metrics compared to other similar countries. Yet, little is known about how leaders affect health care value in the hospitals they run. Second, nonprofit hospitals are particularly important in this context, and it is unclear whether existing findings on executive characteristics can be extrapolated to firms with presumably different incentives and objectives than those which have been previously studied.

@@

    The first prediction of the model is that, all else held equal, hospitals that value profit will have a lower quality than those who value patient welfare in a fee-for-service setting. Second, a hospital's change in quality once pay-for-performance is enacted depends on how much weight the hospital places on profit vs. patient welfare. Particularly, hospitals with higher weight on profit respond more than hospitals with higher weight on patient welfare. Therefore, factors that affect weight $\alpha$ will be correlated with choice of quality, and with how quality changes under new pay-for-performance incentives. In this paper, I test the hypothesis that clinical experience on a hospital's leadership team is a factor that is correlated with weight $\alpha$, and thus affects response to pay-for-performance incentives.

@@

    I construct novel information on individual hospital executives by scraping publicly available tax forms, and merge this to publicly available data sets containing other hospital characteristics: the American Hospital Association Survey (AHA), the Centers for Medicare and Medicaid Services (CMS) Hospital Compare data and Case Mix Index files, and Healthcare Cost Report Information System (HCRIS) data. Ultimately, I employ this hospital-level data spanning from 2010-2014, which contains executive team characteristics, mortality and readmission rates as measures of quality, and other characteristics. In this section, I describe the institutional setting and data collection process. I include more details on the data construction in \ref{appendixdata}.

@@

    Nonprofit hospitals, as with most nonprofits, are typically governed by a board of directors, whose role is complex, but generally is to set broad goals and provide oversight \cite{council_board_roles_2025}. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams. Some hospital executives specialize in health care settings by earning a degree in health care management, or an MBA specific to health care. There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), but medical doctors can also fill other executive positions. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

@@

    For each hospital in the tax form data, I merge to hospital matches in the AHA data based on name and location.\footnote{I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.} For all hospitals that can be matched to a general hospital in the AHA data, I use optical character recognition (OCR) methods to extract the names, titles, and positions of each listed executive from its Tax Form 990. While board members are an important aspect of hospital organization, I do not include them in this analysis as they are beyond the scope of this paper. I extract text spanning 2009-2015, but I retain a much larger sample of hospitals by limiting to hospitals that have text information in 2010-2014. This yields 852 nonprofit hospitals that are (1) matched to AHA data, and (2) contain complete leadership information in each year from 2010-2014.

    Clinical training need not be limited to physicians. However, those are the most common and verifiable clinicians in the data. For the main analysis, I verify clinical training of those who are identified as doctors, and obtain additional information about clinically trained executives by matching names to the National Plan and Provider Enumeration System (NPPES) database of all registered physicians. I present a table of means for various characteristics of doctor executives in the data in Table \ref{doc_sumstats}. In total, there are 813 clinically trained executives. On average, they are 52 years old, and 11\% of clinically trained executives are female. While it is more likely they are a Chief Medical Officer than a Chief Executive Officer, still less than a third are CMOs. The most common specialty is internal medicine. While I cannot perfectly verify nurse experience in the data, in Appendix \ref{app:nurses}, I include executives who are marked as registered nurses in the tax forms. The results are similar to those of the main analysis, which is not surprising given that only 0.5\% of executives have a nursing title.

@@

    As the organization of specific c-suite positions can vary drastically across hospitals, I focus on executive teams as a whole instead of specific positions. It is not uncommon for two hospitals to have the same position but the tasks performed be drastically different, or to have two distinct positions that essentially perform the same tasks. I utilize various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer.

    \subsection{Policies Targeting Readmission and Mortality Rates}\label{sec:hrrp}

    The purpose of this paper is to investigate hospital behavior in the midst of financial incentives that change how hospitals receive payment depending on quality. Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP). HRRP focused on penalizing hospitals with poor quality, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.

@@

    In Table \ref{tab:sumstats_samples_stable}, I present means of relevant variables for several hospital samples used throughout the paper. First, in column (1), are means for hospitals who have a clinical executive at any point in time. Column (2) contains means for hospitals who have a clinical executive for the entire sample period, and column (3) shows means for hospitals who never have a clinical executive in the sample period. The main analysis compares hospitals in columns (2) and (3), and supplemental analyses leverage hospitals in column (1).

@@

    Next, I plot weighted average readmission rates, mortality rates, and patient case mix index across time for different hospital types in Figure \ref{fig:outcomes_graph}. The effective start year of pay-for-performance incentives is shown as a dotted line in 2012. Average readmission rates trend similarly prior to 2012, and all hospital types exhibit a decrease in raw readmission rates after 2012. However, there is a faster rate of decrease for hospitals who have a clinically trained executive at some point, or who never have a clinically trained executive. When considering mortality, there is not a drastic change in 2012 for any hospital type, but hospitals who always have a clinical executive continue on an increasing trend in mortality where the other hospitals types remain more flat.

    \begin{figure}[ht!]
    \centering
        \caption{Main Outcomes Over Time}
        \includegraphics[width=.8\textwidth]{Objects/outcomes_graph.pdf}
        \label{fig:outcomes_graph}
    \end{figure}

@@

    I test empirically whether having clinically trained executives matters in how hospitals respond to the pay-for-performance policies effective in 2012. The quality metrics targeted by the programs are readmission and mortality rates, so I focus on these measures as outcomes. To first analyze a clear control and treatment group, I limit the sample to hospitals who always have clinical executives or never have clinical executives. That is, the hospitals in this analysis do not change their propensity to hire a clinical executive during the sample period.\footnote{An important assumption for this analysis is that this sample selection is not correlated with the program enactment. I investigate this assumption in Section \ref{sec:changes}, and find no evidence that hospitals are more or less likely to change their propensity to hire a clinical executive because of the pay-for-performance programs.} I employ a two-way-fixed-effects strategy as follows:

    \begin{equation}
    \label{eq:clinical}
    y_{ht} = \beta \text{ (clinical training x post 2012)}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
    \end{equation}

    \noindent where the variable $y_{ht}$ is one of the outcome variables capturing readmission or mortality rates discussed in Section \ref{sec:hrrp}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively. The coefficient of interest is $\beta$, capturing the effect of having a clinician on change in quality after 2012, when pay-for-performance policies were rolled out. All hospitals had incentive to improve quality in response to the policy change, as penalties and payments are determined by comparing quality among hospitals. Thus, I leverage the timing of the policy change, rather than the policies themselves.

    One may be concerned about potential differences in quality trends between clinical and non-clinical executive team hospitals leading to a violation in the parallel trends assumption. To account for this, I ultimately implement a synthetic difference-in-differences strategy (\cite{arkhangelsky2021synthetic}), which weights more heavily any non-clinical team hospitals with pre-trends similar to clinical team hospitals in the outcome being considered, essentially guaranteeing no differential pre-trends between always and never clinical teams. Using this strategy, the never-clinical group consists of approximately 240 hospitals, and I show the fairly uniform distribution of their weights in Appendix \ref{app:sumstats}. As a robustness check, I also pre-select the sample of always and never clinical hospitals using the Coarsened Exact Matching (CEM) procedure as described by \citeauthor{azoulay2010superstar} (\citeyear{azoulay2010superstar}). I match based on observable characteristics such as hospital size, number of patients in each condition, and eventual penalty status. For brevity, I only present the synthetic difference-in-differences results here. In Appendix \ref{app:otherestimations}, I show that the results are robust to using standard two-way-fixed-effects, matching, or synthetic control.

    For identification, I assume that under the weighted clinical and non-clinical hospital composition, trends would have continued to be parallel absent the incentive change in 2012. This includes an implicit assumption that no other event occurred in 2012 correlated with both the outcomes and composition of leadership team. Finally, I assume that there is no anticipation of the policy change before 2012. The rules of the policies were announced in October of 2011, so 2012 was the first full year that hospitals could meaningfully respond to the policies.

    The estimated difference in readmission and mortality rates between hospitals with and without clinical executives are shown in Figure \ref{fig:clinicalsynthdid}. Estimate $\beta$ is represented in the difference in slope after 2012 when pay-for-performance incentives began. Panel \ref{fig:read_synth_clinical} shows the estimated difference in readmission rates, and reveals that non-clinical teams decrease readmissions by .3 (standard error .13) more than clinical teams. This is a relatively small magnitude given an average readmission rate of 21. However, \citeauthor{gupta2021impacts} (\citeyear{gupta2021impacts}) finds that hospitals overall decrease readmissions by around 3 percentage points due to the program, so this reveals a meaningful difference in behavior relative to the overall behavior change. In panel \ref{fig:mort_synth_clinical}, I show the estimated difference in mortality for always clinical and never clinical leadership teams after 2012. While the magnitude of the difference is similar to that for readmission rates (relative to the mean), mortality is a noisy measure, and I cannot rule out null effects.

@@

    I present the results of this estimation with readmission rates as the outcome in Figure \ref{fig:cont_clinicalsynthdid}. In panel \ref{fig:belowmed_read_synth_clinical}, I compare executive teams with no clinical training to those with a fraction of clinically trained executives up to the median value, 20\%. With an average treatment effect of .21 (standard error .31), hospitals having this range of clinically trained executives are not driving the main finding. However, in panel \ref{fig:abovemed_read_synth_clinical}, I compare hospitals with no clinically trained executives to hospitals with a fraction of clinically trained executives over the median of 20\%. I find that hospitals without clinically trained executives decrease readmissions by .6 (standard error .14) more than hospitals with clinically trained executives. Thus, while the entire readmission rate response is not driven by highly clinical teams, the effect is exacerbated for these teams that have more clinicians than the median hospital.\footnote{This result holds even when limiting the sample to hospitals with a similar total executive team sizes (Appendix \ref{app:binned})}

@@

    \section{Potential Mechanisms}

    The specific aspects of leadership that link executives to employees to patient outcomes are nearly impossible to observe. As the data on this topic progresses, these granular details of leadership will become more accessible to analyze. For the purpose of this paper, it is important for policymakers to understand why there is a behavioral difference in clinician versus non-clinician leaders, as this might affect the types of incentives they will respond to.

    There are several hypotheses that I test given the data available. First, as the theoretical predictions in Section \ref{sec:theory} state, it is possible that clinicians are less financially driven than non-clinicians, and thus respond less drastically to financial incentives. I test this empirically in Section \ref{sec:forprofit}, comparing the response of nonprofits with different leadership teams to the response of for-profit hospitals on average. Second, I investigate whether clinicians are actively managing hospitals differently or whether they are a signal of underlying preferences of the hospital by leveraging the timing of clinical leaders relative to the incentives in Section \ref{sec:sig_man}. Third, in Section \ref{sec:selection}, I consider whether different types of executives might have different propensities to engage in "gaming" through selective patient practices. In Section \ref{sec:conditions}, I dig deeper into the specific conditions driving the main result, and whether these can help uncover the types of leadership changes made. Similarly, I analyze whether different executives change investment behavior in response to the incentives in Section \ref{sec:investment}. Finally, I consider the altruism versus expertise argument using a heterogeneity analysis of different physician specialties in Section \ref{sec:expertise}.

    \subsection{Financial vs. Patient Driven}\label{sec:forprofit}

    The typical theoretical prediction surrounding quality and pay-for-performance given different valuations of finances vs. patient welfare, shown in Section \ref{sec:theory}, certainly support the idea that, if all else is equal, clinician leaders value patients at baseline more than non-clinician leaders. To investigate this hypothesis empirically, I merge all general for-profit hospitals in the AHA survey to my sample of nonprofit hospitals. I then compare readmission and mortality rates of always clinical executives and never clinical executives to for-profit hospitals after the incentive change.

    While I do not have information on leadership teams of for-profit hospitals, on average, they have revealed through their ownership status that they more heavily weigh profit. Thus, comparing different executive team behavior to for-profit behavior helps reveal how profit driven certain leadership teams are. Thus, I estimate the weighted synthetic difference-in-differences version of

@@

    \subsection{Signaling vs. Managing}\label{sec:sig_man}

@@

    \subsubsection{Analysis of Executive Team Changes} \label{sec:changes}

@@

     \subsubsection{Estimation}\label{sec:decomp}

@@

    \subsection{Patient Selection}\label{sec:selection}

    \subsection{Specific Conditions (rename later)}\label{sec:conditions}

    \subsection{Changes in Investment}\label{sec:investment}

    \subsection{Altruism vs. Expertise}\label{sec:expertise}

@@

    \subsection{Including Nurses in the Sample}\label{app:nurses}

@@

        \begin{figure}[ht!]
     \caption{Effect of Clinical Training on Readmission Rates, Binned Treatment}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho \leq \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_belowmedread_md_nomd_size_synth_graph.pdf}

     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho > \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_abovemedread_md_nomd_size_synth_graph.pdf}

     \end{subfigure}

    \end{figure}

@@

            \begin{figure}[ht!]
     \caption{Effect of Clinical Training on Mortality Rates, Binned Treatment}
     \centering
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho \leq \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_belowmedmort_md_nomd_size_synth_graph.pdf}

     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.45\textwidth}
         \centering
         \caption{$\rho > \eta$}
         \includegraphics[width=\textwidth]{Objects/cont_abovemedmort_md_nomd_size_synth_graph.pdf}

     \end{subfigure}

    \end{figure}
